Decoding the Amplitude-Modulated Part the DCF77 Longwave Time Signal by using the Goertzel Algorithm

### 2024-08-24 paper/main.tex
Explaination of the Goertzel algorithm in matlab finished

            \draw[->] (x.east) -- (sum1.west);
            \draw[->] (sum1.east) -- (y.west);
            \draw[->] (sum2.north) -- (sum1.south);
            \draw[->] (mult1.west) -- (sum2.east);
            \draw[->] (mult2.west) -- (0, -4) -- (sum2.south);
            \draw[->] (4, 0) -- (del1.north);
            \draw[->] (del1.south) -- (del2.north);
            \draw[->] (del2.south) -- (4,-4) -- (mult2.east);
            \draw[->] (4,-2) -- (mult1.east);
            \draw[->] (2, -1.25) -- (mult1.north);
            \draw[->] (2, -3.25) -- (mult2.north);
            \draw[->] (4, -2) -- (st1.west);
            \draw[->] (4, -4) -- (st2.west);
            \node at (2, -1) {$2 \cdot cos(\omega_{g})$};
            \node at (2, -3) {$-1$};
            \node[draw, shape=circle, fill=black, scale=0.3] at (4,0) {};
            \node[draw, shape=circle, fill=black, scale=0.3] at (4,-2) {};
            \node[draw, shape=circle, fill=black, scale=0.3] at (4,-4) {};
        \end{tikzpicture}
    }
    \caption{Block diagram of the Goertzel algorithm without the final calculation of the squared magnitude}
    \label{fig:goertzel_iir}
\end{figure}
\FloatBarrier\noindent
The system takes the input sequence $x[n]$ with length $N$ and is executed exactly $N$ times.
After the execution, the squared magnitude of $f_{a}$, the frequency to be analyzed inside $x[n]$, can be calculated using the values of the two states $Q_{1}$ and $Q_{2}$:
\begin{equation}
    |X(f_{a})|^2 = Q_{1}^2 + Q_{2}^2 - Q_{1} \cdot Q_{2} \cdot 2 \cdot \cos(\omega_{g})
\end{equation}
The constant $2 \cdot cos(\omega_{g})$, often called the "Goertzel coefficient", is calculated before the execution and used to adapt the algorithm to the frequency to be analyzed.
The Value of $\omega_{g}$ itself is calculated based on a integer number $k$ and $N$.
The integer number $k$ is a rounded fraction based on $N$, the frequency to be analyzed $f_{a}$, and the the sampling frequency $f_{s}$:
\begin{equation}
    k = \left\lfloor \frac{N \cdot f_{a}}{f_{s}} \right\rceil
\end{equation}
\begin{equation}
    w_g = \frac{2\pi \cdot k}{N}
\end{equation}

@@

\subsection{Goertzel Algorithm Section}

>>>>

            \draw[->] (x.east) -- (sum1.west);
            \draw[->] (sum1.east) -- (y.west);
            \draw[->] (sum2.north) -- (sum1.south);
            \draw[->] (mult1.west) -- (sum2.east);
            \draw[->] (mult2.west) -- (0, -4) -- (sum2.south);
            \draw[->] (4, 0) -- (del1.north);
            \draw[->] (del1.south) -- (del2.north);
            \draw[->] (del2.south) -- (4,-4) -- (mult2.east);
            \draw[->] (4,-2) -- (mult1.east);
            \draw[->] (2, -1.25) -- (mult1.north);
            \draw[->] (2, -3.25) -- (mult2.north);
            \draw[->] (4, -2) -- (st1.west);
            \draw[->] (4, -4) -- (st2.west);
            \node at (2, -1) {$2 \cdot cos(\omega_{g})$};
            \node at (2, -3) {$-1$};
            \node[draw, shape=circle, fill=black, scale=0.3] at (4,0) {};
            \node[draw, shape=circle, fill=black, scale=0.3] at (4,-2) {};
            \node[draw, shape=circle, fill=black, scale=0.3] at (4,-4) {};
        \end{tikzpicture}
    }
    \caption{Block diagram of the Goertzel algorithm without the final calculation of the squared magnitude}
    \label{fig:goertzel_iir}
\end{figure}
\FloatBarrier\noindent
The system takes the input sequence $x[n]$ with length $N$ and is executed exactly $N$ times.
After the execution, the squared magnitude of $f_{a}$, the frequency to be analyzed inside $x[n]$, can be calculated using the values of the two states $Q_{1}$ and $Q_{2}$:
\begin{equation}
    |X(f_{a})|^2 = Q_{1}^2 + Q_{2}^2 - Q_{1} \cdot Q_{2} \cdot 2 \cdot \cos(\omega_{g})
    \label{eqn:goertzel-mag}
\end{equation}
The constant $2 \cdot cos(\omega_{g})$, often called the "Goertzel coefficient", is calculated before the execution and used to adapt the algorithm to the frequency to be analyzed.
The Value of $\omega_{g}$ itself is calculated based on a integer number $k$ and $N$.
The integer number $k$ is a rounded fraction based on $N$, the frequency to be analyzed $f_{a}$, and the the sampling frequency $f_{s}$:
\begin{equation}
    k = \left\lfloor \frac{N \cdot f_{a}}{f_{s}} \right\rceil
    \label{eqn:goertzel-k}
\end{equation}
\begin{equation}
    w_g = \frac{2\pi \cdot k}{N}
    \label{eqn:goertzel-omega}
\end{equation}

@@

\subsection{Goertzel Algorithm Section}
As described in the chapter explaining the receiver's working principle, the samples are processed in batches whose size is defined by the duration of the Goertzel algorithm's execution.
Before its execution, the parameters needed for the segment-wise execution, such as the segment size, the number of segments, and a time vector containing the exact points in time where results were available (later used for plotting), are calculated based on the
sampling rate $f_{s}$, and the pre-defined duration of analyzation.
The parameters for the Goertzel algorithm itself, $k$, $\omega_{g}$ and $2 \cdot cos(\omega_{g})$ are then calculated by using the formulas \eqref{eqn:goertzel-k} and \eqref{eqn:goertzel-omega}.
\par
Three variables Q\_0, Q\_1, and Q\_2 for the implementation of the Goertzel algorithm's formula (see \eqref{eqn:goertzel} and \eqref{eqn:goertzel-mag}) and a vector for storing the segments' results are initialized.
The segment-wise analyzation is then implemented by using an outer and an inner for-loop.
The outer for-loop iterates over the number of segments, resetting Q\_0, Q\_1, and Q\_2 followed by calculating the index of the first and the last sample of the current segment, providing the limits for the inner for-loop.
Inside the inner for-loop, the formula \eqref{eqn:goertzel} of the IIR-Part is implemented by calculating the new output of the structure and shifting the previous results for the next iteration.
After all samples of the segment were processed, the magnitude is calculated by the implementation of \eqref{eqn:goertzel-mag} using the values stored in Q\_1 and Q\_2 and taking the square root of the result.
The result is then stored inside the prepared vector for the segments' results.
